Thermodynamic fragments for magic states in quantum computation

### 2021-04-09 magic_paper.tex
minor changes

\begin{document}
% Distillation rate vs bound
% Proofs in App. A
% Eq. numbering in last line always
% Sec: App. D
% Sec: General resources
% Sec: Lower bounds
% Sec: Intro
% Sec: Abstract
% Figs consistency (esp. figs 3,4,5 and 1,6) + Figure 0 ?
% Spacing (fig positioning, sections, paragraphs, QED symbols)

@@

There is a natural generalisation, which we call \emph{relative majorization} (also known as $d$--majorization~\cite{Veinott_1971}, or in the context of thermodynamics, thermo-majorization)\footnote{The general form of relative majorization is defined as $(\p,\r) \succ (\q, \r')$ iff there is a stochastic map $A$ such that $A\p = \q$ and $A\r = \r'$.}. For a fixed probability distribution $\r = (r_1, \dots, r_n)$ with positive components, we define majorization relative to $\r$ as follows:
\begin{equation}
\p \succ_{\r} \q,
\end{equation}
if and only if there exists a stochastic map $A$ such that $A\r = \r$ and $A \p = \q$. In order words, the relative majorization pre-order coincides with the image of $\p$ under stochastic maps that have $\r$ as a fixed point. It is readily checked that the original majorization condition between probability distributions corresponds to the case $\r = (1/n, \dots, 1/n)$. The pre-order that results from relative majorization is also easily checked and again corresponds to checking a small number of inequalities.

@@

Given the definition of Lorenz curves, the condition for relative majorization can now be phrased compactly. We have that $\w_1 \succ_{\r} \w_2$ if and only if $L_{\w_1 | \r} (x) \ge L_{\w_2 | \r}(x)$ for all $x \in [0,1]$ (technical discussion and proof can be found in~\cref{app:major}. In other words, the Lorenz curve of $\w_1$ relative to $\r$ is never below the Lorenz curve of $\w_2$ relative to $\r$. Phrased more abstractly, the relative majorization pre-order corresponds to the standard ordering of concave, piece-wise linear functions passing through $(0,0)$ and $(1,1)$. We note that another neat geometric formulation of relative majorization exists, stating that the $L_1$-norm angle between $\w_1$ and $\r$ is greater than or equal to the $L_1$-norm angle between $\w_2$ and $\r$ (see \ddd{[CITE]}\nick{haven't seen this before}\ddd{[`Orthogonality and Disjointness in Spaces of Measures'', by Paul Busch.]} for details).

@@

In this section we apply majorization to magic state distillation within a simple fragment, namely the unital fragment. This corresponds to considering distillation protocols where one is restricted to free operations that are \emph{unital operations}, meaning that $\E(\I/d) = \I/d$. This is a particularly nice set of operations, which include for example all Weyl-covariant channels \ddd{[CITE]}. To make things concrete we shall consider qutrit magic states ($d=3$). This is convenient since this system doesn't have the subtleties of the qubit case, and admits non-trivial structure (e.g. Hamiltonians with two different energy gaps) while being sufficiently simple to analyse in detail.

>>>>

\begin{document}
% Distillation rate vs bound
% Proofs in App. A
% Sec: App. D
% Sec: General resources
% Sec: Lower bounds
% Sec: Intro
% Sec: Abstract
% Figs consistency (esp. figs 3,4,5 and 1,6) + Figure 0 ?
% Eq. numbering in last line always
% British vs US english
% Spacing (fig positioning, sections, paragraphs, QED symbols)

@@

There is a natural generalisation, which we call \emph{relative majorization} (also known as $\bmd$--majorization~\cite{Veinott_1971}, or in the context of thermodynamics, thermo-majorization)\footnote{The general form of relative majorization is defined as $(\p,\r) \succ (\q, \r')$ iff there is a stochastic map $A$ such that $A\r = \r'$ and $A\p = \q$. We retrieve $\bmd$--majorization when $\r = \r'$.}. For a fixed probability distribution $\r = (r_1, \dots, r_n)$ with positive components, we define majorization relative to $\r$ as follows:
\begin{equation}
\p \succ_{\r} \q,
\end{equation}
if and only if there exists a stochastic map $A$ such that $A\r = \r$ and $A \p = \q$. In order words, the relative majorization pre-order coincides with the image of $\p$ under stochastic maps that have $\r$ as a fixed point. It is readily checked that the original majorization condition between probability distributions corresponds to the case $\r = (1/n, \dots, 1/n)$. The pre-order that results from relative majorization is also easily checked and again corresponds to checking a small number of inequalities.

@@

Given the definition of Lorenz curves, the condition for relative majorization can now be phrased compactly. We have that $\w_1 \succ_{\r} \w_2$ if and only if $L_{\w_1 | \r} (x) \ge L_{\w_2 | \r}(x)$ for all $x \in [0,1]$ (technical discussion and proof can be found in~\cref{app:major}. In other words, the Lorenz curve of $\w_1$ relative to $\r$ is never below the Lorenz curve of $\w_2$ relative to $\r$. Phrased more abstractly, the relative majorization pre-order corresponds to the standard ordering of concave, piece-wise linear functions passing through $(0,0)$ and $(1,1)$. We note that another neat geometric formulation of relative majorization exists, stating that the $L_1$-norm angle between $\w_1$ and $\r$ is greater than or equal to the $L_1$-norm angle between $\w_2$ and $\r$ (see~\cite{Busch_1998} for details).

@@

In this section, we apply majorization to magic state distillation within a simple fragment, namely the unital fragment. This encompasses distillation protocols where one is restricted to \emph{unital operations}, meaning that $\E(\I/d) = \I/d$ for all free $\E$. This is a particularly nice set of operations, which include for example all Weyl-covariant channels \ddd{[CITE]}. To make things concrete we shall consider qutrit magic states ($d=3$). This is convenient since qutrit systems don't have the subtleties of the qubit case, and admit non-trivial structure (e.g. Hamiltonians with two different energy gaps), while being sufficiently simple to analyse in detail.
